Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-23 PAPER.tex
made syntax/typo changes from full-read-through up to Section 3.2

We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the \citetalias{ali_et_al2015} analysis and our new analysis) we use $\sim$$8$ hours of data (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

@@

In Section \ref{sec:SiglossOverview}, we discussed signal loss as a form of multiplicative bias to estimates of the signal. Foregrounds are another type of bias, but an additive one instead of multiplicative. Foreground bias is perhaps one of the main factors limiting 21\,cm results, as foreground signals lie $\sim$$4$-$5$ orders of 
magnitude above the cosmological signal. Though there are many techniques proposed for removing foregrounds (see e.g. \citealt{vedantham_et_al2012}; \citealt{parsons_et_al2012a}; \citealt{parsons_et_al2012b}; \citealt{dillon_et_al2013a}; \citealt{wang_et_al2013}; \citealt{parsons_et_al2014}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2015}; \citealt{pober_et_al2016a}; \citealt{trott_et_al2016}), most 
experiments currently remain limited by residuals rather than noise, especially at low $k$.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$ h Mpc$^{-1}$ at $z = 8.4$. The revision of this limit (Kolopanis et al., \textit{submitted}) stems from previously underestimated signal loss and underestimated error bars, both of which we 
address in the following sections.

For our analysis, we use $8.1$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (\citetalias{ali_et_al2015} uses a slightly different RA 
range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ 
MHz ($20$ channels), identical to the analysis in \citetalias{ali_et_al2015}. We note that, besides using only one baseline type instead of the three as in 
\citetalias{ali_et_al2015}, the PAPER-64 dataset that we use in this case study differs from that in \citetalias{ali_et_al2015} only in the analysis. The most significant changes occur in the power spectrum analysis, which is explained in the rest of this paper, but we also note that the applied fringe-rate filter is also slightly different. In \citetalias{ali_et_al2015}, the 
applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

We present our PAPER-64 signal loss investigation in three parts: we first give an overview of our injection framework and illustrate how different power spectrum components affect loss. We next describe our methodology in practice and detail how we map our simulations into a posterior for the EoR signal. Finally, we build off of Section \ref{sec:SiglossOverview} by experimenting with different regularization schemes on PAPER data in order to minimize loss. Throughout each section, we also highlight major differences from the signal loss computation used in \citetalias{ali_et_al2015}, which previously underestimated losses.

@@

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
In short, our method consists of adding an EoR-like signal into the data and then measuring how much of this injected signal would be detectable given any attenuation of this signal by the (lossy) data analysis pipeline.  To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method as in \citetalias{ali_et_al2015}, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from \citetalias{ali_et_al2015} is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in \citetalias{ali_et_al2015} (the dominant reason is explained in the main text in Section \ref{sec:siglossmethod}).}

Suppose that $\textbf{e}$ is a mock injected EoR signal (at some amplitude level), and $\textbf{x}$ is our data. We define $\textbf{r}
$ to be the data plus the EoR signal:

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{\rm out,\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: empirical inverse covariance weighting ($\textbf{R}_{r} = \widehat{\textbf{C}}_{r}^{-1}$) and uniform-weighted ($\textbf{R}_{r} = \textbf{I}$). We will now examine the behavior of this equation in the regimes of very weak and very strong injected signal.

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniform weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a gaussian with zero mean and width $\sqrt{p_e}\sqrt{p_x}$. We also note that in this regime, $\widehat{\textbf{p}}_{r}$ approaches the data level as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than a cosmological signal).  However it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms are small in comparison to the signal-only term. Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the empirical inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation is no longer profoundly negative and can again be modeled as a gaussian distribution of zero mean and width $\sqrt{p_e}\sqrt{p_x}$.

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
the cross-term has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross correlating inverse covariance weighted terms.  This negative component drives down the $P_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.png}
	\caption{Illustration of the power spectrum amplitude as a function of injected power level summed into the 
data. The net output ($P_{out}=P_r$) is shown in black, the individual cross-terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted line indicates 1:1 input to output mapping.  The 
empirical inverse covariance weighted case used in \citealt{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is also the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlation between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{p_e}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the empirical inverse covariance weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small.  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}\footnote{This same appendix also gives a now-prescient warning about signal loss and the dangers of noise and sample variance in inverse covariance.}, to leading order the inversion of a diagonal-dominant matrix like $\widehat{\textbf{C}}_{r}$ (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. Then, when $\widehat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\widehat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\widehat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

{\bf In general:} Another way to phrase the shortcoming of the empirical inverse covariance estimator is that it is not properly normalized. Signal loss due to couplings between the data and its weightings arise because our unnormalized quadratic estimator from Equation \eqref{eq:qhat} ceases to be a quadratic quantity, and instead contains higher order powers of the data. However, the normalization matrix $\mathbf{M}$ is derived assuming that the unnormalized estimator is quadratic in the data. The power spectrum estimate will therefore be incorrectly normalized, which manifests as signal loss. We leave a full analytic solution for $\mathbf{M}$ for future work, since our simulations already capture the full phenomenology of signal loss, and have the added benefit of being more easily generalizable in the face of non-Gaussian systematics.

@@

We now shift our attention towards computing upper limits on the EoR signal for the fringe-rate filtered PAPER-64 dataset in a way that accounts for signal loss. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using empirically estimated inverse covariance weighting 
($\textbf{R} = \widehat{\textbf{C}}^{-1}$), the weighting scheme used in \citetalias{ali_et_al2015} which, with an empirically estimated covariance, leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become:

@@

\begin{equation}
\label{eq:Bayes}
p(P_{\rm in} | P_{\rm out}) \propto {\mathcal{L} (  P_{\rm out} | P_{\rm in})}
  p(P_{\rm in}) ,
\end{equation}

\noindent where $\mathcal{L} $ is the likelihood function defined 
as the distribution of data plus injection ($P_{\rm out}$) given the injection $P_{\rm in}$.  We construct this distribution  
by fixing $P_{\rm in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal 
consistent with this power spectrum. The resulting distribution is normalized such that the sum over $P_{\rm out}$ is unity, and the 
whole process is then repeated for a different value of $P_{\rm in}$.

@@

The implementation details of the injection process require some more detailed explanation. In our code, we add a new realization of EoR to each independent bootstrap of data with the goal of simultaneously capturing variance due to noise and signal loss. To limit computing time we perform $20$ realizations of each $P_{in}$ level. We also run $50$ total EoR injection levels, yielding $P_{\rm in}$ values that range from $\sim$$10^{5}$ mK$^{2}$ (h$^{-1}$ Mpc)$^{3}$ to $\sim$10$^{11}$ mK$^{2}$ (h$^{-1}$ Mpc)$^{3}$, resulting in a total of $1000$ data points on our $P_{\rm in}$ vs. $P_{\rm out}$ grid.

Going forward, we treat every $k$-value separately in order to determine an upper limit on EoR per $k$. For every injection level, we smooth the distribution of $P_{\rm out}$ values using kernel density estimators (\citealt{scott2008kernel}), normalizing each one. Stitching all of them together results in a 2-dimensional smoothed transfer function, shown in Figure \ref{fig:sigloss_transfercurve} for empirical inverse covariance weighting (left) and uniform weighting (right). Although we only show figures for one $k$-value, we note that 
the shape of the transfer curve is nearly identical for all $k$'s. This transfer function is the likelihood function in Bayes' rule, namely $\mathcal{L} (  P_{\rm out} | P_{\rm in})$.  We invoke the Bayes interpretation and re-interpret it as the posterior $P(P_{\rm in}|P_{\rm out})$ where we recall that $P_{\rm out}$ is a model of our data. We then make a horizontal cut across at the data value $P_{\rm out} = \widehat{\textbf{p}}_{x}$ to yield a posterior distribution for the signal. We compute the $95\%$ confidence interval (an upper limit on EoR) from these distributions.

@@

By-eye inspection of the transfer function plot in Figure \ref{fig:sigloss_transfercurve} gives a sense of what the signal loss result should be. The power spectrum value of our data, $
\widehat{\textbf{p}}_{x}$ is marked by the solid gray horizontal lines. From the left plot (empirical inverse covariance weighting), one can eyeball that a data value of $10^{4} \, mK^{2} (h^{-1} Mpc)^{3}$, for example, would map approximately to a 
value of $\sim10^{8} \, mK^{2} (h^{-1} Mpc)^{3}$, implying a signal loss factor of $\sim10000$. For the uniform-weighted case (right plot), we see no loss at a data value of $\sim10^{7} \, mK^{2} (h^{-1} Mpc)^{3}$.

@@

With a signal loss formalism established, we now have the capability of experimenting 
with different weighting options for $\textbf{R}$. Our goal here is to choose a weighting method that successfully down-weights 
foregrounds and systematics in our data without generating large amounts of signal loss. We have found that the balance 
between the two is a delicate one and requires a careful understanding and altering of empirical covariances.

>>>>

We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the \citetalias{ali_et_al2015} analysis and our new analysis) we use $\sim8$ hours of data (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

@@

In Section \ref{sec:SiglossOverview}, we discussed signal loss as a form of multiplicative bias to estimates of the signal. Foregrounds are another type of bias, but an additive one instead of multiplicative. Foreground bias is perhaps one of the main factors limiting 21\,cm results, as foreground signals lie $\sim4$-$5$ orders of 
magnitude above the cosmological signal. Though there are many techniques proposed for removing foregrounds (see e.g. \citealt{vedantham_et_al2012}; \citealt{parsons_et_al2012a}; \citealt{parsons_et_al2012b}; \citealt{dillon_et_al2013a}; \citealt{wang_et_al2013}; \citealt{parsons_et_al2014}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2015}; \citealt{pober_et_al2016a}; \citealt{trott_et_al2016}), most 
experiments currently remain limited by residuals rather than noise, especially at low $k$.

@@

\noindent of $(22.4$ mK$)^{2}$ in the range $0.15 < k < 0.5$ h Mpc$^{-1}$ at $z = 8.4$. The revision of this limit (Kolopanis et al. (\textit{submitted})) stems from previously under-estimated signal loss and under-estimated error bars, both of which we 
address in the following sections.

For our revised analysis, we use $8.1$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (\citetalias{ali_et_al2015} uses a slightly longer RA 
range of $0$-$8.6$ hours; we found that some early LSTs were contaminated). All power spectrum results are produced for a center frequency of 151\,MHz using a width of 10\,MHz ($20$ channels), identical to the analysis in \citetalias{ali_et_al2015}. In the case study in this paper, we only use one baseline type instead of the three as in 
\citetalias{ali_et_al2015}, but Kolopanis et al. (\textit{submitted}) uses all baselines presented in \citetalias{ali_et_al2015}.

The most significant changes from \citetalias{ali_et_al2015} occur in our revised power spectrum analysis, which is explained in the rest of this paper, but we also note that the applied fringe-rate filter is also slightly different. In \citetalias{ali_et_al2015}, the 
applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial 30\,m baseline at 150\,MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

We present our PAPER-64 signal loss investigation in three parts: we first give an overview of our injection framework and illustrate how different power spectrum components affect loss. We next describe our methodology in practice and detail how we map our simulations into a posterior for the EoR signal. Finally, we build off of Section \ref{sec:SiglossOverview} by experimenting with different regularization schemes on PAPER data in order to minimize loss. Throughout each section, we also highlight major differences from the signal loss computation used in \citetalias{ali_et_al2015}, which previously under-estimated losses.

@@

\subsubsection{Signal Loss Methodology} 
\label{sec:siglossmethod}
In short, our method consists of adding an EoR-like signal into the data and then measuring how much of this injected signal would be detectable given any attenuation of this signal by the (lossy) data analysis pipeline.  To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated 21\,cm signal visibilities. Here we use the same method as in \citetalias{ali_et_al2015}, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.\footnote{One 
specific change from \citetalias{ali_et_al2015} is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss, %(by a factor of $\sim$$10$), 
likely due to the use of the fringe-rate filter as a simulator. However, this pipeline difference, while significant, is not the dominant reason why signal loss was under-estimated in \citetalias{ali_et_al2015} (the dominant reason is explained in the main text in Section \ref{sec:siglossmethod}).}

Suppose that $\textbf{e}$ is the mock injected EoR signal (at some amplitude level), and $\textbf{x}$ is our data. We define $\textbf{r}
$ to be the data plus the EoR signal:

@@

\begin{eqnarray}
\label{eq:crossterm}
P_{\rm out,\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: empirically estimated inverse covariance weighting ($\textbf{R}_{r} \equiv \widehat{\textbf{C}}_{r}^{-1}$) and uniform weighting ($\textbf{R}_{r} \equiv \textbf{I}$). We will now examine the behavior of this equation in three different regimes of the injected signal - very weak (left ends of the $P_{\rm in}$ axes in Figure \ref{fig:sigloss_terms}), very strong (right ends), and in between (middle portions).

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a Gaussian with zero mean and width $\sqrt{p_e}\sqrt{p_x}$. We also note that in this regime, $\widehat{\textbf{p}}_{r}$ (black) approaches the data level (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the estimated inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a Gaussian distribution of zero mean and width $\sqrt{p_e}\sqrt{p_x}$.

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
in this regime the cross-term has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross-correlating inverse covariance weighted terms.  This negative component drives down the $P_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.png}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. The net output ($P_{\rm out}=\widehat{\textbf{p}}_r$) is shown in black, the individual terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted black line indicates 1:1 input to output mapping.  The 
empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
horizontal line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{p_e}\sqrt{p_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small.  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}%\footnote{This same appendix also gives a now-prescient warning about signal loss and the dangers of noise and sample variance in inverse covariance.}
, to leading order the inversion of a diagonal-dominant matrix like $\widehat{\textbf{C}}_{r}$ (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. Then, when $\widehat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\widehat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\widehat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

{\bf In general:} Another way to phrase the shortcoming of the inverse covariance estimator is that it is not properly normalized. Signal loss due to couplings between the data and its weightings arise because our unnormalized quadratic estimator from Equation \eqref{eq:qhat} ceases to be a quadratic quantity, and instead contains higher order powers of the data. However, the normalization matrix $\mathbf{M}$ is derived assuming that the unnormalized estimator is quadratic in the data. The power spectrum estimate will therefore be incorrectly normalized, which manifests as signal loss. We leave a full analytic solution for $\mathbf{M}$ for future work, since our simulations already capture the full phenomenology of signal loss and have the added benefit of being more easily generalizable in the face of non-Gaussian systematics.

@@

We now shift our attention towards computing upper limits on the EoR signal for the fringe-rate filtered PAPER-64 dataset in a way that accounts for signal loss. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using empirically estimated inverse covariance weighting 
($\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$), the weighting scheme used in \citetalias{ali_et_al2015} which leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become:

@@

\begin{equation}
\label{eq:Bayes}
p(P_{\rm in} | P_{\rm out}) \propto {\mathcal{L} (  P_{\rm out} | P_{\rm in})}\,p(P_{\rm in}) ,
\end{equation}

\noindent where $\mathcal{L} $ is the likelihood function defined 
as the distribution of data plus injection ($P_{\rm out}$) given the injection $P_{\rm in}$.  We construct this distribution  
by fixing $P_{\rm in}$ and simulating our analysis pipeline for many realizations of the injected EoR signal 
consistent with this power spectrum. The resulting distribution is normalized such that the sum over $P_{\rm out}$ is unity, and the 
whole process is then repeated for a different value of $P_{\rm in}$.

@@

The implementation details of the injection process require some more detailed explanation. In our code, we add a new realization of EoR to each independent bootstrap of data with the goal of simultaneously capturing variance due to noise and signal loss. To limit computing time we perform $20$ realizations of each $P_{\rm in}$ level. We also run $50$ total EoR injection levels, yielding $P_{\rm in}$ values that range from $\sim$$10^{5}$ mK$^{2}$ (h$^{-1}$ Mpc)$^{3}$ to $\sim$10$^{11}$ mK$^{2}$ (h$^{-1}$ Mpc)$^{3}$, resulting in a total of $1000$ data points on our $P_{\rm in}$ vs. $P_{\rm out}$ grid.

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis, and smooth the distribution of $P_{\rm out}$ values using kernel density estimators (\citealt{scott2008kernel}) for each bin (and normalizing them). Stitching all of them together results in a 2-dimensional smoothed transfer function, shown in Figure \ref{fig:sigloss_transfercurve} for both the weighted (left) and unweighted (right) cases. Although we only show figures for one $k$-value, we note that 
the shape of the transfer curve is similar for all $k$'s. This transfer function is the likelihood function in Bayes' rule, namely $\mathcal{L} (  P_{\rm out} | P_{\rm in})$.  We then invoke Bayes' interpretation and re-interpret it as the posterior $p(P_{\rm in}|P_{\rm out})$ where we recall that $P_{\rm out}$ is a model of our data. To do this we make a horizontal cut across at the data value $P_{\rm out} = \widehat{\textbf{p}}_{x}$ to yield a posterior distribution for the signal. We compute the $95\%$ confidence interval (an upper limit on EoR) from these distributions.

@@

By-eye inspection of the transfer function in Figure \ref{fig:sigloss_transfercurve} gives a sense of what the signal loss result should be. The power spectrum value of our data, $
\widehat{\textbf{p}}_{x}$ is marked by the solid gray horizontal lines. From the left plot (empirically estimated inverse covariance weighting), one can eyeball that a data value of $10^{4} \, mK^{2} (h^{-1} Mpc)^{3}$, for example, would map approximately to a 
value of $\sim10^{8} \, mK^{2} (h^{-1} Mpc)^{3}$, implying a signal loss factor of $\sim10^{4}$. For the uniform-weighted case (right plot), we see no loss at a data value of $\sim10^{7} \, mK^{2} (h^{-1} Mpc)^{3}$.

@@

With a signal loss formalism established, we now have the capability of experimenting 
with different weighting options for $\textbf{R}$. Our goal here is to choose a weighting method that successfully down-weights 
foregrounds and systematics in our data without generating large amounts of signal loss as we've seen with the inverse covariance estimator. We have found that the balance 
between the two is a delicate one and requires a careful understanding and altering of empirical covariances.
